---
title: "BAN502_Project_Phase1"
author: "Sayaka Tanaka"
date: "2024-06-01"
output: word_document
---

```{r, include=FALSE}
library(tidyverse) #tidyverse set of packages and functions
library(tidymodels)
library(glmnet) #for Lasso, ridge, and elastic net models 
library(GGally) #create ggcorr and ggpairs plots
library(ggcorrplot) #create an alternative to ggcorr plots
library(gridExtra) #create grids of plots
library(MASS) #access to forward and backward selection algorithms
library(leaps) #best subset selection
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting
library(car) #for calculating the variance inflation factor
library(gridExtra) #used for a little fancy arranging of plots
library(e1071) #often needed for various statistical tasks
library(ROCR) #for threshold selction
library(skimr)
library(mice)
library(UpSetR)
library(VIM) #visualizing missingness
library(naniar)
```

```{r}
train = read_csv("train.csv")

str(train)
summary(train)
nrow(train)
```

Rename
```{r}
train = train %>%
  rename(
    a0 = attribute_0,
    a1 = attribute_1,
    a2 = attribute_2,
    a3 = attribute_3,
    m0 = measurement_0,
    m1 = measurement_1,
    m2 = measurement_2,
    m3 = measurement_3,
    m4 = measurement_4,
    m5 = measurement_5,
    m6 = measurement_6,
    m7 = measurement_7,
    m8 = measurement_8,
    m9 = measurement_9,
    m10 = measurement_10,
    m11 = measurement_11,
    m12 = measurement_12,
    m13 = measurement_13,
    m14 = measurement_14,
    m15 = measurement_15,
    m16 = measurement_16,
    m17 = measurement_17
  )
```

Convert all character variables to factors  
```{r}
train = train %>% mutate_if(is.character,as_factor)
```

Convert attribute_02 and 03 to factor
```{r}
train = train %>% mutate(a2 = as_factor(a2))
train = train %>% mutate(a3 = as_factor(a3))
```

Make a new numeric variable "failure_num" Yes=1, No = 0
```{r}
train$failure_num = ifelse(train$failure == "Yes", 1, 0)
```

Missing data?
```{r}
skim(train)
```

```{r}
gg_miss_var(train)
```

```{r}
gg_miss_fct(x = train, fct = failure)
```

```{r}
vis_miss(train)
```

#### Drop missing data  
##### Since the missing data is fairly evenly distributed between the 'yes' and 'no' failure categories and not concentrated in either, I decided to remove rows with missing data.  
```{r}
train = train %>% drop_na()
skim(train)
str(train)
```
#### How many of them failed?
```{r}
table(train$failure)
```

```{r}
# bar chart for "failure" variable
table(train$failure)
ggplot(train, aes(x = failure, fill = failure)) + geom_bar(stat = "count") + labs(title = "Distribution of Failure", x = "Failure")+theme_bw()

```

```{r}
failure_ratio = prop.table(table(train$failure))

print(failure_ratio)

failure_ratio_df = as.data.frame(failure_ratio)
colnames(failure_ratio_df) = c("failure", "ratio")

ggplot(failure_ratio_df, aes(x = failure, y = ratio, fill = failure)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Ratio of Failure", x = "Failure", y = "Ratio") + 
  scale_y_continuous(labels = scales::percent)+theme_bw()
```

#### Visualization of factor Variables vs failure  
```{r}
p1 = ggplot(train, aes(x = product_code, fill = failure)) + geom_bar(position = "fill") +theme_bw()+xlab("product_code")
p2 = ggplot(train, aes(x = a0, fill = failure)) + geom_bar(position = "fill") +theme_bw()+xlab("attribute_0")
p3 = ggplot(train, aes(x = a1, fill = failure)) + geom_bar(position = "fill") +theme_bw()+xlab("attribute_1")
p4 = ggplot(train, aes(x = a2, fill = failure)) + geom_bar(position = "fill") +theme_bw()+ xlab("attribute_2")
p5 = ggplot(train, aes(x = a3, fill = failure)) + geom_bar(position = "fill") +theme_bw()+ xlab("attribute_3")
grid.arrange(p1,p2,p3,p4,p5)
```

```{r}
p6 = ggplot(train, aes(x = failure, y = loading,fill=failure)) + geom_boxplot() +theme_bw()+ylab("loading")
p7 = ggplot(train, aes(x = failure, y = m0,fill=failure)) + geom_boxplot() +theme_bw()+ylab("Measurement_0")
p8 = ggplot(train, aes(x = failure, y = m1,fill=failure)) + geom_boxplot() +theme_bw()+ylab("Measurement_1")
grid.arrange(p6,p7,p8, ncol = 2)
```

```{r}
p9 = ggplot(train, aes(x = failure, y = m2,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_2")
p10 = ggplot(train, aes(x = failure, y = m3,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_3")
p11 = ggplot(train, aes(x = failure, y = m4,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_4")
p12 = ggplot(train, aes(x = failure, y = m5,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_5")
grid.arrange(p9,p10,p11,p12, ncol = 2)
```

```{r}
p13 = ggplot(train, aes(x = failure, y = m6,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_6")
p14 = ggplot(train, aes(x = failure, y = m7,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_7")
p15 = ggplot(train, aes(x = failure, y = m8,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_8")
p16 = ggplot(train, aes(x = failure, y = m9,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_9")
grid.arrange(p13,p14,p15,p16, ncol = 2)
```

```{r}
p17 = ggplot(train, aes(x = failure, y = m10,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_10")
p18 = ggplot(train, aes(x = failure, y = m11,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_11")
p19 = ggplot(train, aes(x = failure, y = m12,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_12")
p20 = ggplot(train, aes(x = failure, y = m13,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_13")
grid.arrange(p17,p18,p19,p20, ncol = 2)
```
 
```{r}
p21 = ggplot(train, aes(x = failure, y = m14,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_14")
p22 = ggplot(train, aes(x = failure, y = m15,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_15")
p23 = ggplot(train, aes(x = failure, y = m16,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_16")
p24 = ggplot(train, aes(x = failure, y = m17,fill=failure)) + geom_boxplot()+theme_bw()+ylab("Measurement_17")
grid.arrange(p21,p22,p23,p24, ncol = 2)
```


```{r}
ggpairs(train,columns = c( "failure","loading","product_code", "a0", "a1"))
```

```{r}
ggpairs(train,columns = c( "failure","a2","a3","m0","m1"))
```


```{r}
ggpairs(train,columns = c( "failure","m2","m3","m4","m5"))
```

```{r}
ggpairs(train,columns = c( "failure","m6","m7","m8","m9"))
```

```{r}
ggpairs(train,columns = c( "failure","m10","m11","m12","m13"))
```

```{r}
ggpairs(train,columns = c( "failure","m14","m15","m16","m17"))
```


## Use "failure_num" variable for analysis
### ‘loading’ variable shows the largest correlation coefficient of 0.1.

```{r}
ggcorr(train,label = TRUE,label_round = 1)


corr = round(cor(train[, c(3, 8:25,27)]), 1)
ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE)

```

```{r}
ggpairs(train,columns = c( "failure_num","loading"))
```



```{r}
ggpairs(train,columns = c( "failure_num","loading","product_code", "a0", "a1"))
```

```{r}
ggpairs(train,columns = c( "failure_num","a2","a3","m0","m1"))
```

```{r}
ggpairs(train,columns = c( "failure_num","m2","m3","m4","m5"))
```


```{r}
ggpairs(train,columns = c( "failure_num","m6","m7","m8","m9"))
```

```{r}
ggpairs(train,columns = c( "failure_num","m10","m11","m12","m13"))
```

```{r}
ggpairs(train,columns = c( "failure_num","m14","m15","m16","m17"))
```
#### Visualization of numeric variables vs failure_num  
```{r}
p25 = ggplot(train, aes(x = product_code, y = failure_num)) + geom_boxplot() +theme_bw()+xlab("product_code")
p26 = ggplot(train, aes(x = a0, y = failure_num)) +  geom_boxplot() +theme_bw()+xlab("attribute_0")
p27 = ggplot(train, aes(x = a1, y = failure_num)) +  geom_boxplot() +theme_bw()+ xlab("attribute_1")
p28 = ggplot(train, aes(x = a2, y = failure_num)) + geom_boxplot() +theme_bw()+ xlab("attribute_2")
p29 = ggplot(train, aes(x = a3, y = failure_num)) +  geom_boxplot() +theme_bw()+ xlab("attribute_3")
grid.arrange(p25,p26,p27,p28,p29)
```

#### Visualization of numeric variables vs failure_num  
```{r}
p30 = ggplot(train, aes(x = failure_num, y = loading)) + geom_point() +theme_bw()+ylab("loading")
p31 = ggplot(train, aes(x = failure_num, y = m0)) + geom_point() +theme_bw()+ylab("Measurement_0")
p32 = ggplot(train, aes(x = failure_num, y = m1)) + geom_point() +theme_bw()+ylab("Measurement_1")
grid.arrange(p30,p31,p32, ncol = 2)
```
  
```{r}
p33 = ggplot(train, aes(x = failure_num, y = m2)) + geom_point()+theme_bw()+ylab("Measurement_2")
p34 = ggplot(train, aes(x = failure_num, y = m3)) + geom_point()+theme_bw()+ylab("Measurement_3")
p35 = ggplot(train, aes(x = failure_num, y = m4)) + geom_point()+theme_bw()+ylab("Measurement_4")
p36 = ggplot(train, aes(x = failure_num, y = m5)) + geom_point()+theme_bw()+ylab("Measurement_5")
grid.arrange(p33,p34,p35,p36, ncol = 2)
```
  
```{r}
p37 = ggplot(train, aes(x = failure_num, y = m6)) + geom_point()+theme_bw()+ylab("Measurement_6")
p38 = ggplot(train, aes(x = failure_num, y = m7)) + geom_point()+theme_bw()+ylab("Measurement_7")
p39 = ggplot(train, aes(x = failure_num, y = m8)) + geom_point()+theme_bw()+ylab("Measurement_8")
p40 = ggplot(train, aes(x = failure_num, y = m9)) + geom_point()+theme_bw()+ylab("Measurement_9")
grid.arrange(p37,p38,p39,p40, ncol = 2)
```
 
```{r}
p41 = ggplot(train, aes(x = failure_num, y = m10)) + geom_point()+theme_bw()+ylab("Measurement_10")
p42 = ggplot(train, aes(x = failure_num, y = m11)) + geom_point()+theme_bw()+ylab("Measurement_11")
p43 = ggplot(train, aes(x = failure_num, y = m12)) + geom_point()+theme_bw()+ylab("Measurement_12")
p44 = ggplot(train, aes(x = failure_num, y = m13)) + geom_point()+theme_bw()+ylab("Measurement_13")
grid.arrange(p41,p42,p43,p44, ncol = 2)
```


```{r}
p45 = ggplot(train, aes(x = failure_num, y = m14)) + geom_point()+theme_bw()+ylab("Measurement_14")
p46 = ggplot(train, aes(x = failure_num, y = m15)) + geom_point()+theme_bw()+ylab("Measurement_15")
p47 = ggplot(train, aes(x = failure_num, y = m16)) + geom_point()+theme_bw()+ylab("Measurement_16")
p48 = ggplot(train, aes(x = failure_num, y = m17)) + geom_point()+theme_bw()+ylab("Measurement_17")
grid.arrange(p45,p46,p47,p48, ncol = 2)
```